Multiple deprivation index using small area estimation methods: an application for the adult population in Colombia

### 2022-03-10 Paper_MPI_CEPAL/main.tex
Final changes abstract
Description

\author[1]{Andrés Gutiérrez-Rojas}
\author[2]{Alejandra Arias-Salazars}
\author[1]{Diego Lemus-Polanía}
\author[3]{Natalia Rojas-Perilla}
\author[1]{Xavier Mancero}
\affil[1]{Economic Commission for Latin America and the Caribbean}
\affil[2]{Freie Universität Berlin}
\affil[3]{United Arab Emirates University}

@@

Based on the new ``Multiple deprivation index" (MDI) for Latin American countries produced by the Economic Commission for Latin America and the Caribbean, this paper shows the case study of Colombia to obtain estimates of this multidimensional index in small domains. This country has a recent population census (2018) providing most of the information required to compute the MDI at municipality level. However,  two of the indicators required to compute the index are not available in the census, namely: ``Employment-related income" and ``Lack of social security". Before computing the index, small area estimation (SAE) methods are implemented to obtain unit-level estimates for those specific indicators. Since two different SAE procedures are carried out, a parametric bootstrap algorithm is proposed to provide respective uncertainty measures.

@@

The MDI is defined as: 
\textcolor{red}{briefly describe how it is computed.}

\begin{table}[ht]
\begin{center}
\small{
	\caption{Dimensions, indicators of the MDI and data availability in the census}
	\centering
		\begin{tabular}{llcc}
		\hline
				Dimension &Indicator&Available in census &Target population  \\
				\hline
	Housing &Housing materials & Yes & Adults and seniors\\
	& Overcrowding & Yes & Adults and seniors \\
			\hline
	Basic services &  Information communication & Yes & Adults and seniors \\
	 & technologies (ICTs) & &\\
	& Drinking water & Yes & Adults and seniors \\
	& Sanitation & Yes & Adults and seniors \\
	&Energy (electricity) & Yes & Adults and seniors \\
			\hline
	Education & Unfinished education & Yes &Adults \\
	             & Illiteracy& Yes & Seniors \\
	             		\hline
	  Employment, social  & Pension & Yes & Seniors \\
	  security and income & Employment-related income & No &Adults \\
	   & Unemployment, precarious  & No & Adults\\
	   & employment and out of the &&\\
	   &labor force due to housework&&\\

>>>>

\author[1]{Andrés Gutiérrez-Rojas}
\author[2]{Alejandra Arias-Salazar}
\author[1]{Diego Lemus-Polanía}
\author[3]{Natalia Rojas-Perilla}
\author[1]{Xavier Mancero}
\affil[1]{Economic Commission for Latin America and the Caribbean}
\affil[2]{Freie Universität Berlin}
\affil[3]{United Arab Emirates University}

@@

Based on the new ``Multiple deprivation index" (MDI) for Latin American countries produced by the Economic Commission for Latin America and the Caribbean, this paper shows the case study of Colombia to obtain estimates of this multidimensional index in small domains. This country has a recent population census (2018) providing most of the information required to compute the MDI at municipality level. However,  two of the indicators required to compute the index are not available in the census, namely: ``Employment-related income" and ``Lack of social security". Therefore, small area estimation (SAE) methods are implemented to obtain unit-level estimates for those specific indicators. Since two different SAE procedures are carried out, a parametric bootstrap algorithm is proposed to provide respective uncertainty measures.

@@

The MDI, similarly as the Global Multdimensional Poverty Index (\cite{Alkire}), is defined as  the product of two measures: the incidence of multidimensional poverty and the intensity of poverty:

\begin{equation*}
    \text{MDI} = \text{H} \cdot \text{A}
\end{equation*}

where H is the headcount ratio or incidence of multidimentional poverty and A is the intensity of poverty. The Headcount ratio is defined as:

\begin{equation*}
    \text{H} = \frac{q}{n}
\end{equation*}

with $q$ the number of people who are multidimensionally poor and $n$ the total population. The intensity of poverty A is defined as:

\begin{equation*}
    \text{A} = \frac{\sum_{1}^{q} s_i}{q}
\end{equation*}

with $s_i$ the deprivation score that the $i_th$ multidimentionally poor person experiences.

For this case study, the indicators and dimensions considered are explained in table 1. Each dimension has same weight (1/4) and one individual is defined as poor if he or she obtains 40\% of total deprivations. Notice that this index is based on individual information, i.e. quantity of deprivations by individual. In case one member of the household is defined multidimensionally poor, all members will received the same status.

\begin{table}[ht]
\begin{center}
\small{
	\caption{Dimensions, indicators of the MDI and data availability in the census}
	\centering
\begin{tabular}{llccc}
				\hline
				Dimension &Indicator& Weight &Available in census &Target population  \\
				\hline
				Housing &Poor housing materials & 1/8 & Yes & Adults and seniors\\
				& Overcrowding & 1/8 &Yes & Adults and seniors \\
				\hline
				Basic services &  Lack of information communication & 1/16 &Yes & Adults and seniors \\
				& technologies (ICTs) & &\\
				& Lack of drinking water & 1/16&Yes & Adults and seniors \\
				& Lack of sanitation &1/16 & Yes & Adults and seniors \\
				& Lack of energy (electricity) & 1/16&Yes & Adults and seniors \\
				\hline
				Education & Unfinished education & 1/4 &Yes &Adults \\
				& Illiteracy& 1/4 &Yes & Seniors \\
				\hline
				Employment, social  & No pension & 1/4 &Yes & Seniors \\
				security and income & Insufficient employment-related  &1/8 & \textbf{No} &Adults \\
				&income&&&\\
				& Unemployment, precarious  &1/8 & \textbf{No} & Adults\\
				& employment and out of the &&&\\
				&labor force due to housework&&&\\
